Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-12 prop_letter.tex
prop_letter.tex \section{Born} and \section{Gleason} Updating...

In order to find a function~$\iota$ satisfying Eq.~(\ref{eq:iota-Inclusion}),
we decompose it into two functions: its left-end and right-end~$\left[\iotal\left(x\right),\iotar\left(x\right)\right]=\iota\left(x\right)$.
By Eq.~(\ref{eq:iota-Inclusion}), the left-end function~$\iotal:\left[0,1\right]\rightarrow\left[0,1\right]$
is Wright-convex, i.e., $\iotal\left(x_{0}+x_{1}+x_{2}\right)+\iotal\left(x_{2}\right)\ge\iotal\left(x_{0}+x_{2}\right)+\iotal\left(x_{1}+x_{2}\right)$,
and the right-end function~$\iotar:\left[0,1\right]\rightarrow\left[0,1\right]$
is Wright-concave, i.e., $\iotar\left(x_{0}+x_{1}+x_{2}\right)+\iotar\left(x_{2}\right)\le\iotar\left(x_{0}+x_{2}\right)+\iotar\left(x_{1}+x_{2}\right)$~\citep{Wright1954,RobertsVarberg1973,PecaricTong1992,Brzdek_2014}.
It is well-known that a Wright-convex function is midpoint-convex
or Jensen-convex~\citep{Klee_1956,Kenyon_1956,RobertsVarberg1973,PecaricTong1992},
i.e., we have
\begin{equation}
\frac{\iotal\left(x_{0}\right)+\iotal\left(x_{1}\right)}{2}\ge\iotal\left(\frac{x_{0}+x_{1}}{2}\right)\,.
\end{equation}
Together with the fact that $\iotal$ maps to a bounded interval~$\left[0,1\right]$,
the left-end function~$\iotal$ must be continuous on the unit open
interval~$\left(0,1\right)$~\citep{Jensen_1906,RobertsVarberg1973}.
Therefore, either $\iota$ maps every number in the same interval,
or the number of intervals that $\iota$ maps to must be uncountable.

\begin{thm}\label{thm:convex-uncountable}Let $\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} $.
For any function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
described in Thm.~\ref{thm:iota-statements}, either $\mathscr{I}$
is $\mathscr{I}_{0}$ or $\mathscr{I}$ contains uncountably many
intervals.\end{thm}

@@

\yutsung{The other idea: We never proved or find an counterexample
that a non-convex QIVPM has empty core, and its expectation value
is not the same as defined from core... If it is true, the convex
condition may not be necessary, but this might be hard to prove...}

@@

In other words, if the measurement intervals are too imprecise, the
results we get might contradict to each other, and exclude any possible
state from being consistent with the measurement results. In contrast
with the QIVPM defined in Table~\ref{tab:non-Born-QIVPM}, if a QIVPM~$\iota\circ\mu_{\rho}^{\mathrm{B}}$
is induced by a conventional quantum probability measure~$\mu_{\rho}^{\mathrm{B}}$
and a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$ considered
in Sec.~\ref{sec:Born}, there is always a state consistent with
it because $\rho\in\coreBorn\left(\iota\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.
Moreover, the collection of states consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to $\rho$ as the intervals get narrower and
narrower. To state this result, we introduce a few technical notions.
First, the norm of a pure state is defined as usual $||\psi||=\sqrt{\ip{\psi}{\psi}}$.
Then, the norm of the difference of two states~$\rho$ and $\rho'$
is just the usual operator norm
\begin{equation}
\left\Vert \rho-\rho'\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert \left(\rho-\rho'\right)\ket{\psi}\right\Vert \,,
\end{equation}
which is also commonly known as $L^{2}$ norm or spectral norm~\citep{GolubVanLoan1996,Foucart2012}.

\begin{thm}[A finite-precision extension of weak Gleason's theorem
for Born QIVPMs]\label{thm:Finite-precision-Gleason}Consider a three-dimensional
Hilbert space, and an interval-valued function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
with the maximum length of intervals in $\mathscr{I}$ is $\delta$.
Given any state~$\rho$, if a state $\rho'\in\coreBorn\left(\iota\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
then the norm of their difference is $\left\Vert \rho-\rho'\right\Vert \le\delta$.\end{thm}

Given a state~$\rho'$ consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$,
we have $\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\in\iota\left(\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right)$
for any 1-dimensional projector~$\proj{\psi}$. Since the maximum
length of the intervals in $\mathscr{I}$ is $\delta$, it is also
the upper bound of the difference $\left|\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right|=\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$.
Since $\rho-\rho'$ is Hermitian, $\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$is
the maximum absolute value of the eigenvalues of $\rho-\rho'$~\citep{544199},
and equal to $\left\Vert \rho-\rho'\right\Vert $~\citep{GolubVanLoan1996,Foucart2012}.
This implies $\left\Vert \rho-\rho'\right\Vert \le\delta$ by definition.

>>>>

In order to find a function~$\iota$ satisfying Eq.~(\ref{eq:iota-Inclusion}),
we decompose it into two functions: its left-end and right-end~$\left[\iotal\left(x\right),\iotar\left(x\right)\right]=\iota\left(x\right)$.
By Eq.~(\ref{eq:iota-Inclusion}), the left-end function~$\iotal:\left[0,1\right]\rightarrow\left[0,1\right]$
is Wright-convex, i.e., $\iotal\left(x_{0}+x_{1}+x_{2}\right)+\iotal\left(x_{2}\right)\ge\iotal\left(x_{0}+x_{2}\right)+\iotal\left(x_{1}+x_{2}\right)$,
and the right-end function~$\iotar:\left[0,1\right]\rightarrow\left[0,1\right]$
is Wright-concave, i.e., $\iotar\left(x_{0}+x_{1}+x_{2}\right)+\iotar\left(x_{2}\right)\le\iotar\left(x_{0}+x_{2}\right)+\iotar\left(x_{1}+x_{2}\right)$~\citep{Wright1954,RobertsVarberg1973,PecaricTong1992,Brzdek_2014}.
It is well-known that a Wright-convex function is midpoint-convex
or Jensen-convex~\citep{Klee_1956,Kenyon_1956,RobertsVarberg1973,PecaricTong1992},
i.e., we have
\begin{equation}
\frac{\iotal\left(x_{0}\right)+\iotal\left(x_{1}\right)}{2}\ge\iotal\left(\frac{x_{0}+x_{1}}{2}\right)\,.
\end{equation}
Together with the fact that $\iotal$ maps to a bounded interval~$\left[0,1\right]$,
the left-end function~$\iotal$ must be continuous on the unit open
interval~$\left(0,1\right)$~\citep{Jensen_1906,RobertsVarberg1973}.
Therefore, either $\iota$ maps every number to the same interval,
or the number of intervals that $\iota$ maps to must be uncountable.

\begin{thm}\label{thm:convex-uncountable}Let 
\[
\mathscr{I}_{0}=\left\{ \necess,\imposs,\unknown\right\} \,.
\]
For any function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
described in Thm.~\ref{thm:iota-statements}, either $\mathscr{I}=\mathscr{I}_{0}$
or $\mathscr{I}$ contains uncountably many intervals.\end{thm}

@@

\yutsung{The other idea: We never proved a non-convex QIVPM always
have a non-empty core or find an counterexample that a non-convex
QIVPM has empty core. Moreover, we have no idea the relation between
its expectation value and the Choquet integral... Depending the on
situation, the convex condition may not be necessary, but this might
be hard to prove...}

@@

In other words, if the measurement intervals are too imprecise, the
results we get might contradict to each other, and exclude any possible
state from being consistent with the measurement results. In contrast
with the QIVPM defined in Table~\ref{tab:non-Born-QIVPM}, if a QIVPM~$\iota\circ\mu_{\rho}^{\mathrm{B}}$
is induced by a conventional quantum probability measure~$\mu_{\rho}^{\mathrm{B}}$
and a function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$ considered
in Sec.~\ref{sec:Born}, there is always a state consistent with
it because $\rho\in\coreBorn\left(\iota\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.
Moreover, the collection of states consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$
gets closer and closer to $\rho$ as the intervals get narrower and
narrower. To state this result, we introduce a few technical notions.
First, the norm of a pure state is defined as usual $||\psi||=\sqrt{\ip{\psi}{\psi}}$.
Then, the norm of the difference of two states~$\rho$ and $\rho'$
is just the usual operator norm
\begin{equation}
\left\Vert \rho-\rho'\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert \left(\rho-\rho'\right)\ket{\psi}\right\Vert \,,
\end{equation}
which is also commonly known as $2$-norm or spectral norm~\citep{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.

\begin{thm}[A finite-precision extension of weak Gleason's theorem]\label{thm:Finite-precision-Gleason}Consider
a three-dimensional Hilbert space, and an interval-valued function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
with the maximum length of intervals $\delta$. Given any state~$\rho$,
if a state $\rho'\in\coreBorn\left(\iota\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
then the norm of their difference is $\left\Vert \rho-\rho'\right\Vert \le\delta$.\end{thm}

Given a state~$\rho'$ consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$,
we have $\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\in\iota\left(\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right)$
for any 1-dimensional projector~$\proj{\psi}$. Since the maximum
length of the intervals in $\mathscr{I}$ is $\delta$, it is also
the upper bound of the difference $\left|\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right|=\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$.
Since $\rho-\rho'$ is Hermitian, $\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$
is the maximum absolute value of the eigenvalues of $\rho-\rho'$~\citep{544199},
and equal to $\left\Vert \rho-\rho'\right\Vert $~\citep{GolubVanLoan1996,Foucart2012}.
Hence, $\left\Vert \rho-\rho'\right\Vert \le\delta$.
